---
title: "Working with Zonation results"
author: "Joona Lehtomäki"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Zonation results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo = FALSE, message = FALSE} 
hasData <- requireNamespace("zdat", quietly = TRUE) 
if (!hasData) {
  knitr::opts_chunk$set(eval = FALSE)
  msg <- paste("Note: Examples in this vignette require that the", 
               "`zdat` package be installed. The system", 
               "currently running this vignette does not have that package", 
               "installed, so code examples will not be evaluated.")
  msg <- paste(strwrap(msg), collapse = "\n") 
  message(msg)
} 
```

```{r setup, echo=FALSE}
options(width = 120)
```

## 1. Introduction to working with Zonation results

First, let's set up the scene. For the examples below to work, make sure you have installed [`zdat`](https://github.com/cbig/zdat) package. If not, do the following:

```{r install-zdat, eval=FALSE}
install.packages('zdat', repos='https://jlehtoma.github.io/drat/', 
                 type='source')
```

Now we start working with `zonator`.

```{r setup-objects, message=FALSE}
library(zonator)

# Start by creating a project using the tutorial data
setup.dir <- system.file("extdata/basic", package = "zdat") 
tutorial.project <- load_zproject(setup.dir)

# Get a specific variant, 01_core_area_zonation
variant.caz <- get_variant(tutorial.project, "01")

# Let's also rename the features and groups while we're at it
featurenames(variant.caz) <- c("Koala", "Masked.owl", "Powerful.owl", 
                             "Tiger.quoll", "Sooty.owl", "Squirrel.glider",
                             "Yellow-bellied.glider")

groupnames(variant.caz) <- c("1" = "mammals", "2" = "owls")
```

Each variant can either already have results (i.e. Zonation run has completed) 
or then not. Which results exactly are available can be determined by using 
method `has_results()`.

```{r has-results, message=FALSE} 
has_results(variant.caz)
```

Value `TRUE` indicates that Zonation has already been run and results are 
available and `FALSE` indicates that results are not available.

Results of a given variant - if present - can be extracted using `results()`.
The resulting object will be an instance of `Zresults` class.

```{r get-results}
results.caz <- results(variant.caz)
```

Object `results.caz` can in turn be queried with various dedicated methods
described in more detail below.

----

## 2. Performance curves

### 2.1 Working with performance data

Raw curves data (with or without groups) can be extracted from the variant or 
results using `curves()`. Optional argument `groups` defines whether feature or 
group specific curves are returned. If you don't need every feature or aggregate 
statistic, you can also get the data for particular features.


```{r curves}
# Get feature-specific curves
feature.curves.caz <- curves(results.caz)
head(feature.curves.caz)

# Note! Using curves() variant.caz will work as well
feature.curves.caz <- curves(variant.caz)
head(feature.curves.caz)

# Get just 2 species
duo.curves.caz <- curves(results.caz, cols = c("Koala", "Tiger.quoll"))
head(duo.curves.caz)
```

Columns are exactly the same as in Zonation curves output file. Note also that 
the actual feature columns are named according to feature names in the 
`Zvariant` object these results were derived from. Thus whatever names you 
assigned to features prior to extracting the results will be also seen here.

Let's see  what the group curves data looks like as a whole or for individual 
groups:

```{r grp-curves}
# Get group-specific curves
feature.grp.curves.caz <- curves(results.caz, groups = TRUE)
head(feature.grp.curves.caz)

# Get all the aggregate stats for owls
owls.grp.curves.caz <- curves(results.caz, groups = TRUE, 
                              cols = c("min.owls", "mean.owls", "max.owls",
                                     "w.mean.owls", "ext2.owls"))
head(owls.grp.curves.caz)
```

Often a particular range of cell removal process is more interesting than the 
whole range. `curves()` has arguments `lost.lower` and `lost.upper` to define 
a specific range:

```{r curves-range}

# Get performance values for 2 species for a range between 60% and 75% of 
# landscape lost 
duo.curves.caz.range <- curves(results.caz, cols = c("Koala", "Tiger.quoll"),
                               lost.lower = 0.6, lost.upper = 0.75)
head(duo.curves.caz.range)
tail(duo.curves.caz.range)
```

Sometimes you might also be interested in performance values of features at 
specific levels of landscape protection. For example, how much of the original
occurrence levels are remaining when 25%, 50%, and 75% of landscape is lost.
Method `performance()` can give you exactly this information.

```{r perfomance}
# Define the levels of landscape lost
lost.levels <- c(0.25, 0.50, 0.75)
# Notice that performance operates with Zresults-object
performance(results.caz, lost.levels)

# Same can be done for groups
performance(results.caz, lost.levels, groups = TRUE)
```

**Pro tip!** If you're a heavy user of Hadley Wickham's excellent 
[ggplot2](http://ggplot2.org/) package you might a lot of time 
[reshaping](http://www.statmethods.net/management/reshape.html) your data. 
Method `performance()` also has an optional argument `melted` which if set TRUE 
will make the method return performance data in a melted format and thus readily 
suitable for ggplot2-plotting (see below for a plotting example).

```{r perfomance-melted}
# Define the levels of landscape lost
lost.levels <- c(0.25, 0.50, 0.75)
# Notice that performance operates with Zresults-object
performance(results.caz, lost.levels, melted = TRUE)
```

### 2.2 Plotting performance curves

To make plotting examples more interesting, let's use also another variant that
uses Zonation's masking capability.

```{r mask-variant}
# Variant: 05_hierarchical_removal_mask
variant.mask <- get_variant(tutorial.project, "05")

featurenames(variant.mask) <- c("Koala", "Masked.owl", "Powerful.owl", 
                             "Tiger.quoll", "Sooty.owl", "Squirrel.glider",
                             "Yellow-bellied.glider")
labels <- c("mammals", "owls")
names(labels) <- c(1, 2)
groupnames(variant.mask) <- labels

results.mask <- results(variant.mask)
feature.curves.mask <- curves(results.mask)
feature.grp.curves.mask <- curves(results.mask, groups = TRUE)
```

Performance data (curves) can also be easily plotted, feature- or group-wise:

```{r curves-plotting-1, fig.width=7, fig.height=5}
plot(feature.curves.caz)
```

Plotting-methods for `ZCurvesDataFrame` and `ZGroupCurvesDataFrame` classes 
include a couple of convenience arguments as well. By default, the scale on x-axis
ranges from 0 to 1 and has a label "Prop. of landscape lost". The other way of 
presenting the x-axis is to have it range from 1 to 0 and label it "Proportion
of landscape under conservation". This behavior can be controlled by using the
argument `invert.x` for `plot-method`. The title of the plot can be set using
`main`-argument.

```{r curves-plotting-2, fig.width=7, fig.height=5}
plot(feature.curves.caz, invert.x = TRUE, main = "All features")
```

Argument `subs` can have a character vector of feature names to be plotted, i.e.
a subset of all features. Different overall statistics can also be plotted. 
Statistics supported are: `mean`, `min`, and `w.mean` for feature-specific 
curves. Each of these is a logical argument that can be set to `TRUE`/`FALSE`.
They correspond to "Average proportion remaining over all features", 
"Minimum proportion remaining over all features", and
"Weighted average proportion remaining over all features" respectively. 

```{r curves-plotting-3, fig.width=7, fig.height=5}
plot(feature.curves.caz, invert.x = TRUE, main = "Just few features", 
     subs = c('Yellow-bellied.glider', 'Sooty.owl'), mean = TRUE)
```

Group-specific curves can be plotted in similar fashion. Note that plotting 
group curves also respect the group names you may have set.

```{r curves-plotting-4, fig.width=7, fig.height=5}
plot(feature.grp.curves.caz, invert.x = TRUE, main = "All groups")
```

Subsetting works the same as with individual features.

```{r curves-plotting-5, fig.width=7, fig.height=5}
plot(feature.grp.curves.caz, invert.x = TRUE, main = "All groups", 
     subs = c('owls'))
```

Both feature- and group-specific curve plotting methods also have an argument
called `monochrome`. Setting this argument to `TRUE` will use only monochrome
colours which may be useful e.g. for publication figures.

```{r curves-plotting-6, fig.width=7, fig.height=5}
plot(feature.grp.curves.caz, invert.x = TRUE, main = "All groups", 
     monochrome = TRUE)
```

Group plotting also has a set of statistics which can be plotted, but the 
behavior is slightly different to that of features. What is plotted by default 
are group-specific mean values. Other statistics are `min`, `max`, and `w.mean`
which correspond to "Minimum proportion remaining for a feature within group",
"Maxmimum proportion remaining for a feature within group", and 
"Weighted average proportion remaining for group". Note that `mean` will always
be `TRUE` unless it's explicitly set to `FALSE`.

```{r curves-plotting-7, fig.width=7, fig.height=5}
plot(feature.grp.curves.mask, invert.x = TRUE, main = "All groups", 
     mean = TRUE, min = TRUE, max = TRUE)
```

### 2.3 Plotting performance levels

If you want to summarize how much each feature has retained it's occurrence 
level in a given stage of the cell removal process, you can easily created 
bar plots out of selected performance levels. For example, if you'd like to 
know how much of the original occurrence level each of species in our tutorial 
data has left when 10% percent of the landscape is under conservation, you 
could do the following:

```{r perfomance-hist, fig.width=7, fig.height=5}
library(ggplot2)
# NOTE: we're using melted=TRUE here
occr.remaining <- performance(results.caz, c(0.9, 0.95, 0.98), melted = TRUE)
ggplot(occr.remaining, aes(x = factor(feature), y = perf.levels,
                           fill = factor(pr.lost))) + 
  geom_bar(position = "dodge", stat = "identity")

```

---

## 3. Raster results

### 3.1. Priority rank rasters

One of the main outputs of a Zonation analysis is the priority rank rasters
which gives each cell in the landscape a value [0, 1] indicating the final 
priority rank (0 = the lowest, 1 = the highest). Rank raster is part of 
`Zresults` object just as the performance curves above. Method `rank_raster()`
can be used to get the rank raster object. The color scheme used is really up 
to the user, but having a consistent and expressive color scheme for all the
results helps. `zlegends()` can be used for ready-to-go color schemes that can 
used when plotting the raster objects.

```{r rank-raster-1, fig.width=7, fig.height=5}
library(raster)
rank.caz <- rank_raster(results.caz)
# Let's use a spectral color scheme designed for Zonation rank maps
leg <- zlegend("spectral")
plot(rank.caz, breaks = leg$values, col = leg$colors)
```

You can of course get all rank rasters individually from your project's 
variants, but an alternative approach is to call `rank_rasters()` (note
the plural in method name!) directly on your project object. This will result
in a `RasterStack` object containing all the rank rasters as respective layers.
If you want to plot all rank rasters in your project for comparison, you could 
do the following:

```{r rank-raster-2, warning=FALSE, fig.width=7, fig.height=20}
# RasterStack is returned
all.ranks <- rank_rasters(tutorial.project)
plot(all.ranks, breaks = leg$values, col = leg$colors, nc = 1)
```

Naturally, all methods defined for a `RasterStack` object are available. For example, checking 
the distributions of individual stack layers (i.e. rank rasters) and the 
correlation between the values of layers can be done using `pairs()`. Following
example examines correlation between variants 1-3.

```{r rank-raster-3, fig.width=7, fig.height=5}
pairs(stack(all.ranks[[1:3]]), method = "kendall")
```

Values are sampled from the whole distribution of values to increase
execution speed. Note also that we are using `kendall` as method for 
correlation.

Masking is another useful analysis tool. In the following example we are using
the mask file `mask_rs.tif` shipping with Zonation basic tutorial in order to
inspect the distribution of priorities within a subset of the landscape (i.e.
protected areas, PAs). In the following example, note that method 
`rank_raster()` has been implemented for class `Zvariant`. In other words, you 
can call it on both your variant or results object. `plot_hist()` accepts 
two rasters as arguments: first is the rank raster and the second one is a mask
raster. Histogram is then plotted based on the values of the rank raster within
the mask raster.

```{r rank-raster-4, fig.width=7, fig.height=5}
pas.file <- system.file("extdata/data", "mask_rs.tif", 
                        package = "zdat")
pas <- raster(pas.file)
# Re-class zeros as NA
pas[pas == 0] <- NA
# Plot the histogram of rank.caz within mask raster
plot_hist(x = rank.caz, pas, add.median = TRUE)
```

[rasterVis](http://oscarperpinan.github.io/rastervis/) is another useful 
package for plotting raster data. For example, `levelplot()` can be used to 
plot data and show a marginal plot of of a function across each coordinate
(here using mean): 

```{r rank-rasterVis-1, fig.width=7, fig.height=5, message=FALSE}
library(rasterVis)
# Note that breaks in the following are different from the previous maps
levelplot(rank.caz, FUN.margin = mean, 
          par.settings = rasterTheme(region = leg$colors))
```
