<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Create zonator project based on existing data}
-->

```{r setup, echo=FALSE}
options(width=120)
```

## 1. Create zonator project based on existing data

The following examples will use 
[the basic Zonation tutorial data](https://github.com/cbig/zonation-tutorial/tree/master/basic)
that is distributed with zonator package. This includes also the results of the
tutorial runs so there is no need to run the variants in order to inspect the
results. In case you *do* want to rerun the variants and you have Zonation
installed in your system do so by running the following code. Otherwise just
skip this code section.

```{r run-variant-1, eval=(1:6), message=FALSE}
library(zonator)

setup.dir <- system.file("extdata/tutorial/basic", package="zonator")

# Get all the bat-files
bat.files <- list.files(setup.dir, "\\.bat$", full.names=TRUE)

# Run all the runs
#for (bat.file in bat.files) {
#  run_bat(bat.file)
#}

```

You can create a new zonator project based on existing results by doing the 
following:

```{r creat-zproject, message=FALSE}
tutorial.project <- create_zproject(setup.dir)
```

**Pro tip!** In case you have a complex Zonation project and want to keep track
which files are being read in while creating zproject, you can do 
`create_zproject(setup.dir, debug=TRUE)` in order to enable logging of file
reading sequence.

zonator also includes a utility function `opendir()` which takes a zproject
object as an argument and opens the file system folder containing the setup
files:

```{r opendir, eval=FALSE}
opendir(tutorial.project)
```

## 2. Working with variants

Individual variants (i.e. runs) can be extracted form the project using an 
index number. `nvariants()` will tell how many variants are included in the 
project.

```{r extract-variant}
nvariants(tutorial.project)
variant.1 <- get_variant(tutorial.project, 1)
```

Using an index number such as `1` is one option, but you can also use the name
of the variant. By default, zonator will assign the name of bat-file used to run
the run as a name, without the ".bat" extensions of course. `names()` will print 
the names of all the variants. Name can also be used to extract a variant.

```{r variants-by-name}
names(tutorial.project)
variant.caz <- get_variant(tutorial.project, "01_core_area_zonation")
```

Each variant object is an instance of class `Zvariant` and have a suite of 
useful methods for dealing with data parsed from various Zonation input files.

### 2.1 spp data

Zonation spp-file is one of the mandatory input files that always needs to be in
place and thus all variants have one. When a new `Zvariant` instance is created
the associated spp file is automatically parsed into it. All the spp data 
(with group code column if available) can be retrieved using `sppdata()`:

```{r sppdata}
sppdata(variant.caz)
```

You can check the number of features in spp file by using method `nfeatures()`:

```{r nfeatures}
nfeatures(variant.caz)
```

`Zvariant` objects have a couple of other convenience functions for quickly 
accessing columns in the spp file. For example, `sppweights()` can be used to 
extract the weight column in spp file:

```{r sppdata-weights}
# Note that all biodiversity features (species) have an equal weight of 1
sppweights(variant.caz)
```

Feature names from the spp file/data can be accessed directly by using method
`featurenames()`.

```{r sppdata-names-2}
featurenames(variant.caz)
```

The generated names are not necessarily very informative and can be changed to
new values. Remember that the names need to be valid data frame column names
(zontor will try to fix these even if you don't).

```{r sppdata-names-1}
featurenames(variant.caz) <- c("Koala", "Masked.owl", "Powerful.owl", 
                             "Tiger.quoll", "Sooty.owl", "Squirrel.glider",
                             "Yellow-bellied.glider")
featurenames(variant.caz)
# Or all the spp data
sppdata(variant.caz)
```

### 2.2 Groups

Notice that data frames returned by `sppdata()` in previous examples already had
a column called "group". This is because all tutorial variants have groups 
enabled by default. If a variant doesn't use groups, then this column will be
missing.

Group identities in Zonation input file are coded with integer values. Method
`groups()` will return just this integer vector:

```{r groups}
groups(variant.caz)
```

Groups can also have more informative names attached to them by using method
`groupnames()`. Even if you haven't names the groups used (Zonation does not
have a concept of named groups), generic group names "group1", "group2" etc. 
will automatically be created for `Zvariant` objects with groups enabled.

```{r group-names-1}
# By default, generic group names are used
groupnames(variant.caz)
```

The format for setting (mapping) group names is strict and involves a named 
character vector in which (column) names correspond to integer group codes and 
character elements the group names to be assigned:

```{r group-names-2}
# Construct a group name mapping using a named character vector
groupnames(variant.caz) <- c("1"="mammals", "2"="owls")
groupnames(variant.caz)
```

Now group 1 is labeled "mammals" and group 2 is labeled "owls". Note that 
`sppdata()` has an optional argument `group.names` that can be set to `TRUE`
if group names are preferable to group codes.

```{r sppdata-group-names}
sppdata(variant.caz, group.names=TRUE)
```

In addition to changing the group names, you can change the groups themselves 
as well. In the previous examples we had two groups named "mammals" (group code
 = 1) and "owls" (group code = 2). Belonging to either one of these groups is 
 controlled by the group ID code (an integer number) that can be accessed by
 `groups()`. As was shown above, doing `groups(variant.caz)` will yield a result
 of ``r groups(variant.caz)``.
 
Method `groups()` can be used to set the group ID codes and thus the grouping. 
Let's say you want to split owls into two sub-groups called "big.owls" (group
code = 2) containing the Sooty owl and the Masked Owl and "small.owls" (group 
code = 3) containing the Powerful owl. Take again a look at the spp data:

```{r changing-groups-1}
sppdata(variant.caz)
```

What needs to be done is to change the group ID for the Powerful owl so that 
the group ID vector changes from ``r groups(variant.caz)`` to 
`1, 2, 3, 1, 2, 1, 1`. So do the following:

```{r changing-groups-2}
groups(variant.caz) <- c(1, 2, 3, 1, 2, 1, 1)
# Check the group ID codes
groups(variant.caz)
```

Note that after you change the group ID codes all the group names you may have
assigned get lost and are again replaced by generic "group1", "group2" etc. You
will have to manually set the group names again:

```{r changing-groups-3}
# Using generic group names after the group IDs have been changed
groupnames(variant.caz)
# Construct a new group names mapping
groupnames(variant.caz) <- c("1"="mammals", "2"="big.owls", "3"="small.owls")
groupnames(variant.caz)
```

or check the whole spp data with group names:

```{r changing-groups-4}
sppdata(variant.caz, group.names=TRUE)
```
